## Splines

```{r}
Sys.setlocale("LC_ALL", "C")
#install.packages("gam")
require(ISLR)
require(splines)
names(Wage)
attach(Wage)
```

## Cubic Splines / Fixed Knot Regression Splines

* splines are more flexible
* spline is a cubic polynomial
* knots are places of discontinuity
* splines are more local than polynomial
* nice way of fitting flexible functions
* bs() generates the basis

```{r}
fit=lm(wage~bs(age,knots=c(25,40,60)),data=Wage)
fit
plot(age,wage,col="darkgrey")
lines(age.grid,predict(fit,list(age=age.grid)),col="darkgreen",lwd=2)
abline(v=c(25,40,60),lty=2,col="darkgreen")
```

## Smoothing Splines

* 16 degrees of freedom
* too wiggly in this case

```{r}
fit2 = smooth.spline(age,wage,df=16)
lines(fit2,col="red",lwd=2)
```

* use loocv to determine the smoothing parameter automatically
* will get warning about many ties in x, but this is acceptable
* cv will use 6.79 degrees of freedom

```{r}
fit3 = smooth.spline(age,wage,cv=T)
lines(fit3,col="purple",lwd=2)
fit3$df

```

## Generalized Additive Models

* call gam package, fit a generalized additive model
* tell it the response is wage
* tell it we want a smooth term in age with 4 deg of freedom
* tell it we want a smooth term in year with 4 deg of freedom
* education is a factor var, so we have it in as a linear term
* s() is the smoothing spline function of gam 
* age and year will be fit jointly
* par(mfrow=c(1,3)) sets plot layout to 3 columns

```{r fig.width=10,  fig.height=5  }
require(gam)
gam1 = gam(wage~s(age,df=4)+s(year,df=4)+education,data=Wage)
par(mfrow=c(1,3))
plot(gam1,se=T)
```

* gam also works with logistic regression
* leave off the std errors
* plot the contributions to the logit, of the probability, each as a separate function

```{r}
gam2 = gam(I(wage>250)~s(age,df=4)+s(year,df=4)+education,data=Wage,family = binomial)
plot(gam2)
```

* lets see if we need a non-linear term for the variable year
* do anova to test gam2 and gam2a
* chisq p value is 0.8242, says we dont need this nonlinear term for year, a linear term will be fine

```{r}
gam2a = gam(I(wage>250)~s(age,df=4)+year+education,data=Wage,family = binomial)
anova(gam2a,gam2,test="Chisq")
```

* use the plot method to plot models fit by lm and glm
* fit a linear model with wage as response
* instead of using gam, we'll fit lm
* generate a spline basis using ns() (natural splines)
* instead of knots, specify defgrees of freedom
* since lm is not of class gam, we write out plot.gam() to use that plot method
* plots a fitted function, including std errors

```{r  fig.width=10 , fig.height=5}
par(mfrow=c(1,3))
lm1 = lm(wage~ns(age,df=4)+ns(year,df=4)+education,data=Wage)
plot.gam(lm1,se=T)

```






