## Linear Support Vector Machine

* generate some data, in 2 dimensions
* plot the data and color code it, depending on if y is +1 or -1

```{r}
set.seed(10111)
x = matrix(rnorm(40),20,2)
y=rep(c(-1,1),c(10,10))
x[y==1,] = x[y==1,]+1
plot(x,col=y+3,pch=19)
```

* load package e1071
* create dataframe from x and y (coerced into a factor variable)
* columns from x are unpacked into x1 and x2 in dataframe

```{r}
library(e1071)
dat = data.frame(x,y=as.factor(y))
summary(dat)
```

* fit svm (a linear) classifier
* scale=F means we are not standardizing the variables
* cost=10, this is the tuning parameter
* show summary of the model
* number of support vectors (points close to boundary or on the wrong side of boundary) is 6

```{r}
svmfit = svm(y~. , data=dat , kernel="linear" , cost=10 , scale=F)
svmfit
```

* plot svm 
* plot is bad because it puts X2 on horiz. and X1 on vert. axis


```{r}
plot(svmfit,dat)
```


* make function that makes a grid
  + apply range to each variable in x
  + make a grid of length n using seq()
  + expand.grid makes the lattice

```{r}
make.grid=function(x,n=75)
{
  grange=apply(x,2,range)
  x1=seq(from=grange[1,1],to=grange[2,1],length=n)
  x2=seq(from=grange[1,2],to=grange[2,2],length=n)
  expand.grid(X1=x1,X2=x2)
}
```

* apply function
* 5625 points on grid
* predict from svmfit at the values on xgrid (new data)

```{r}
xgrid=make.grid(x)
dim(xgrid) # 5625 points on grid
xgrid[1:10,]
ygrid=predict(svmfit,xgrid)

```

* plot all points in xgrid, color according to prediction
* see the decision boundary
* each point is a point on the lattice
* identify all of the original points, close to the boundary
* circle points that are instrumental to making the decision boundary

```{r}
plot(xgrid,col=c("red","blue")[as.numeric(ygrid)],pch=20,cex=.2)
points(x,col=y+3,pch=19)
points(x[svmfit$index,],pch=5,cex=2)
```

* svm in R is not very user friendly
* we don't ge the coefficients used to describe the linear function
* check chap 12 of ESL to do this
* extract beta and beta0

```{r}
beta=drop(t(svmfit$coefs)%*%x[svmfit$index,])
beta0=svmfit$rho
```

* re-plot points, with the decision boundary
* some of the support points are exactly on the margin
* some are inside the margin

```{r}
plot(xgrid,col=c("red","blue")[as.numeric(ygrid)],pch=20,cex=.2)
points(x,col=y+3,pch=19)
points(x[svmfit$index,],pch=5,cex=2)
abline(beta0/beta[2],-beta[1]/beta[2])
abline((beta0-1)/beta[2],-beta[1]/beta[2],lty=2)
abline((beta0+1)/beta[2],-beta[1]/beta[2],lty=2)
```

