## Non Linear Models

```{r}
Sys.setlocale("LC_ALL", "C")
require(ISLR)
names(Wage)
attach(Wage)
```

## Polynomial with Linear Regression

* fit a 4th degree polynomial
* degree 1 and 2 are very significant, 3 pretty significant
* the function poly() generates orthogonal polynomials

```{r}
fit1 = lm(wage~poly(age,4),data=Wage)
summary(fit1)
```

* make a grid from min to mix of the age variable
* make predictions (predictions$fit), with standard errors (predictions$se.fit)

```{r}
age.grid = seq(from=min(age),to=max(age))
predictions = predict(fit1,newdata=list(age=age.grid),se=T)
```

* make standard error bands
* plot age vs wage
* add line for fit and matlines for std errors
* can use a package to do this instead

```{r fig.width=7 , fig.height=6}
se.bands = cbind(predictions$fit+2*predictions$se.fit,predictions$fit-2*predictions$se.fit)
plot(age,wage,col="darkgrey")
lines(age.grid,predictions$fit,lwd=2,col="blue")
matlines(age.grid,se.bands,col="blue",lty=2)
```

* now re-do the same fit, but without using poly() 
* I() is an indentity function, protects what's inside
* p values for fit2 are different, different basis
* but the fited values are the same

```{r}
fit2 = lm(wage~age+I(age^2)+I(age^3)+I(age^4),data=Wage)
summary(fit2)
plot(fitted(fit1),fitted(fit2))
```

* if you have other variables in the model, use anova to test significance of models
* anova says age and age squared models are significant, others not

```{r}
fita = lm(wage~education,data=Wage)
fitb = lm(wage~education+age,data=Wage)
fitc = lm(wage~education+poly(age,2),data=Wage)
fitd = lm(wage~education+poly(age,3),data=Wage)
anova(fita,fitb,fitc,fitd)  
```

## Polynomial with Logistic Regression

* create a high wage earners binary variable
* make predictions with std errors
* make std error bands, 0, -2 std errors and 2 std errors, 3 col matrix
* look at first 5 rows of the matrix


```{r}
fit3 = glm(I(wage>250) ~ poly(age,3), data=Wage, family=binomial)
summary(fit3)
predictions2 = predict(fit3,list(age=age.grid),se=T)
se.bands = predictions2$fit + cbind(fit=0,lower=-2*predictions2$se.fit,upper=2*predictions2$se.fit)
se.bands[1:5,]
```

* make plot, estimate of fitted probability in solid blue
* estimate of 95% conf interval
* add points to show where data occurs

```{r}
prob.bands = exp(se.bands)/(1+exp(se.bands))
matplot(age.grid,prob.bands,col="blue",lwd=c(2,1,1),lty=c(1,2,2),type='l',ylim=c(0,0.1))
points(jitter(age),I(wage>250)/10,pch="l",cex=0.5)
```










