Intro
=================

Using R Markdown for this project
See "Markdown Quick Reference" for more info

```{r}
Sys.setlocale("LC_ALL", "C")
#install.packages("leaps")
require(ISLR)
require(leaps)
```

Our response variable is Salary

```{r}
summary(Hitters$Salary)
```

Remove missing values from dataframe

```{r}
Hitters = na.omit(Hitters)
with(Hitters,sum(is.na(Hitters)))
```

## Best subset regression

* Agressive, looks at all possible subsets.
* Produces a sequence of models, best subset for each size.
* By default it goes up to subset of size 8.
* Don't use this method with over 10-20 predictors

```{r}
regfit.full = regsubsets(Salary~. , data=Hitters)
summary(regfit.full)
```

Re-run with 19 variables and save summary to a variable

```{r}
ncol(Hitters)-1 # how many max variables used?
regfit.full = regsubsets(Salary~. , data=Hitters , nvmax = 19)
regfit.summary = summary(regfit.full)
regfit.summary
```

Plot Cp in a "map plot"", black means in, white means out

```{r}
plot(regfit.full,scale="Cp")
```

We want to pick the model with the lowest Cp

```{r}
names(regfit.summary)
regfit.summary$cp
which.min(regfit.summary$cp)
regfit.summary$cp[10]
```

Plot the Cp statistic vs number of variables
Color the point with lowest Cp
pch means plotting character (plotting symbol) used

```{r}
plot( regfit.summary$cp , xlab="Num of Variables", ylab="Cp" )
points(10,regfit.summary$cp[10],pch=20,col="red")
```

Output a coefficient vector for the 10 variables in the model

```{r}
coef(regfit.full,10)
```





