## Forward Stepwise Selection

```{r}
Sys.setlocale("LC_ALL", "C")
#install.packages("leaps")
require(ISLR)
require(leaps)
```

* Greedy algorithm
* Produces a nested sequence of models

```{r}
regfit.fwd = regsubsets(Salary ~ . , data = Hitters , nvmax=19 , method = "forward")
regfit.fwd.summary = summary(regfit.fwd)
regfit.fwd.summary
```

Plot the schematic

```{r}
plot(regfit.fwd,scale="Cp")
```

## Model selection using a validation set

* Remove missing values from dataframe
* Create a train set from a random index of 180/263 numbers
* 


```{r}
Hitters = na.omit(Hitters)
with(Hitters,sum(is.na(Hitters)))
dim(Hitters)
set.seed(1)
numbers = sample(seq(263),180,replace = F)
numbers
train = Hitters[numbers,] #index by rows from numbers vector
regfit.fwd = regsubsets(Salary ~ . , data = train ,nvmax=19,method="forward")
```

* make predictions on test set
* make a vector with 19 empty slots, one for each variable
* make an x matrix corresponding to test set


```{r}
test = Hitters[-numbers,]
val.errors = rep(NA,19) # make 19 nulls
x.test = model.matrix(Salary~. , data=test) # test matrix
```

* for each model
    + coef to extract coefficients
    + for each size
    + index the columns by the names on coefficient vector
    + do a matrix, multiplied by coefficient vector
    + compute the mean squared error
    
```{r}

for(i in 1:19)
{
  coefi = coef(regfit.fwd,id=i)
  pred = x.test[,names(coefi)]%*%coefi
  val.errors[i]=mean((test$Salary-pred)^2)
}

```

Make a plot

```{r}
plot(sqrt(val.errors),ylab="Root MSE",ylim=c(300,400),pch=19,type="b")
points(sqrt(regfit.fwd$rss[-1]/180),col="blue",pch=19,type="b")
legend("topright",legend=c("Training","Validation"),col=c("blue","black"),pch=19)
```

Write our own predict regsubset function

```{r}
predict.regsubsets=function(object,newdata,id,...)
{
  form = as.formula(object$call[[2]])
  mat=model.matrix(form,newdata)
  coefi = coef(object,id=id)
  mat[,names(coefi)%*%coefi]
}
```


